Supporting Information : Complex and informative dependencies propogate throughout ecology and can be revealed using copulas

### 2019-05-13 Paper.Rmd
about to try to switch the order of some sup mat figs, committing in case somethig goes wrong

Our causal and consequential hypotheses imply, or rely on, an "observational
hypothesis" (O in Fig. \ref{MasteFigure}), which we here frame: datasets in 
ecology will commonly have copula structure
distinct from that of a Gaussian/normal distribution (here called 
non-normal copula structure); they will commonly show tail dependence 
distinct from that of a normal distribution; in particular they will
show asymmetric tail dependence.

@@

For the model we examined, causal hypothesis D (Fig. \ref{MasterFigure}) was 
correct that asymmetric tail dependence in evolutionary changes 
can produce similarly asymmetric tail dependence in character distributions across
phylogeny tips (Appendix \ref{SM-evol_mech} fro details). This topic is revisited in the Discussion.

>>>>

Our causal and consequential hypotheses imply, or rely on, an "observational
hypothesis" (O in Fig. \ref{MasterFigure}), which we here frame: datasets in 
ecology will commonly have copula structure
distinct from that of a Gaussian/normal distribution (here called 
non-normal copula structure); they will commonly show tail dependence 
distinct from that of a normal distribution; in particular they will
show asymmetric tail dependence.

@@

For the model we examined, causal hypothesis D (Fig. \ref{MasterFigure}) was 
correct that asymmetric tail dependence in evolutionary changes 
can produce similarly asymmetric tail dependence in character distributions across
phylogeny tips (Appendix \ref{SM-evol_mech} for details). This topic is revisited in the Discussion.

### 2019-05-13 SupportingInformation.Rmd
about to try to switch the order of some sup mat figs, committing in case somethig goes wrong

# From causes to consequences for green spruce aphid abundance, details \label{sect:greenspruce}
Winter temperature for year $t$ was for December of year $t-1$ to March of 
year $t$. Data, which pertained to the same locations and years as the aphid data,
were obtained and processed exactly as in [@Sheppard2016]. Copulas
with winter temperature are in Fig. \ref{fig:greenspruce}. Results of analyses of these
copulas that paralleled the analyses applied to bivariate ecological datasets 
(see section on methods for the observational hypotheses) are in Table \ref{tab:greenspruce}.

@@

      if(IndepTestRes<0.05 && tauval>0){
        plot(vi,vj,type='p',col=rgb(0,0,0,0.3),pch=19,xlim=c(0,1),ylim=c(0,1),
             xlab=names(data_aphid_count_with_temp)[sp],ylab="winter temp.",cex.lab=1.5)
        mtext(paste0("Loc = ",loc," , p = ",round(IndepTestRes,3)),side = 3, line=0.15, adj=0.5)
      }else if(IndepTestRes<0.05 && tauval<0){
        plot(vi,vj,type='p',col=rgb(0,1,0,0.3),pch=19,xlim=c(0,1),ylim=c(0,1),
             xlab=names(data_aphid_count_with_temp)[sp],ylab="winter temp.",cex.lab=1.5)
        mtext(paste0("Location = ",loc," , p = ",round(IndepTestRes,3)),side = 3, line=0.15, adj=0.5)
      }else{
        plot(-1,0,xlim=c(0,1),ylim=c(0,1),xlab=names(data_aphid_count_with_temp)[sp],ylab="winter temp.",cex.lab=1.5)
        text(0.5,0.5,"Indep.",adj=c(0.5,.5),cex=2)
        mtext(paste0("Location = ",loc," , p = ",round(IndepTestRes,3)),side = 3, line=0.15, adj=0.5)
      }

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/aphid_count_winter_temp_results/copulaplot_greenspruceaphidcount_temp.pdf}
\caption[Copula plot for green spruce aphid abundance data with winter temperature]{Copula plot for green spruce 
aphid abundance data with winter temperature for each selected locations. \label{fig:greenspruce}}
\end{center}
\end{figure}

@@

res<-ms_npa_RES_greenspruceaphid_count_temp$sp10_Green_spruce_aphid$ms_summary_allloc
#row.names(res)<-NULL
#res<-cbind(c(2:11),res)
colnames(res)<-c("p-value from indep. test","bestfit copula",
                 "bestfit copula AICw","Normal Copula AICw","p-value from CvM stats","p value from KS stats","LT dependence","UT dependence","(LT - UT) dependence","cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$")
res<-res[,-1]
knitr::kable(res,"latex", booktabs = T, linesep = "",digits=3, escape=F,longtable=T,
             caption.short="Summary results for copula made with time-series of green spruce aphid abundance data and winter temperature for each selected locations",
      caption = "Summary results for copula made with time-series of green spruce aphid abundance data and winter temperature for each selected locations, tail dependence were calculated using model averaged AIC-weights, NA represents copula was independent at given location as p-value from independence test was beyond the threshold 0.05. Last three columns indicate non-parametric statistics computed on $(0,0.5)$ and $(0.5,1)$ bounds for lower and upper statistics, respectively.\\label{tab:greenspruce}") %>% column_spec(1:12,width="1.2 cm")#%>%
#kable_styling(latex_options = c("scale_down"))
```

>>>>

# From causes to consequences for green spruce aphid abundance, details \label{sect:greenspruce}
Winter temperature for year $t$ was for December of year $t-1$ to March of 
year $t$. Temperature data, which pertained to the same locations and years as the aphid data,
were obtained and processed exactly as in @Sheppard2016. Copulas
describing the relationship between green spruce aphid abundance and winter 
temperature are in Fig. \ref{fig:greenspruce}. Results of analyses of these
copulas that paralleled the analyses applied to bivariate ecological datasets 
(see section on methods for the observational hypotheses) are in Table \ref{tab:greenspruce}.

@@

      if(IndepTestRes<0.05 && tauval>0){
        plot(vi,vj,type='p',col=rgb(0,0,0,0.3),pch=19,xlim=c(0,1),ylim=c(0,1),
             xlab="Aphid abundance",ylab="winter temp.",cex.lab=1.5)
        mtext(paste0("Location = ",loc," , p = ",round(IndepTestRes,3)),side = 3, line=0.15, adj=0.5)
      }else if(IndepTestRes<0.05 && tauval<0){
        plot(vi,vj,type='p',col=rgb(0,1,0,0.3),pch=19,xlim=c(0,1),ylim=c(0,1),
             xlab="Aphid abundance",ylab="winter temp.",cex.lab=1.5)
        mtext(paste0("Location = ",loc," , p = ",round(IndepTestRes,3)),side = 3, line=0.15, adj=0.5)
      }else{
        plot(-1,0,xlim=c(0,1),ylim=c(0,1),xlab=names(data_aphid_count_with_temp)[sp],ylab="winter temp.",cex.lab=1.5)
        text(0.5,0.5,"Indep.",adj=c(0.5,.5),cex=2)
        mtext(paste0("Location = ",loc," , p = ",round(IndepTestRes,3)),side = 3, line=0.15, adj=0.5)
      }

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/aphid_count_winter_temp_results/copulaplot_greenspruceaphidcount_temp.pdf}
\caption[Copula plots for winter temperature versus green spruce aphid abundance]{Copula plots 
for winter temperature versus green spruce 
aphid abundance for each of the 10 sampling locations. The $p$-values in the header are from the test of independence (see section on methods for the observational hypothesis).\label{fig:greenspruce}}
\end{center}
\end{figure}

@@

res<-ms_npa_RES_greenspruceaphid_count_temp$sp10_Green_spruce_aphid$ms_summary_allloc
#row.names(res)<-NULL
#res<-cbind(c(2:11),res)
colnames(res)<-c("p-value, independence test","Best-fit copula",
                 "Best-fit copula AICw","N-copula AICw","p, CvM","p, KS","Model-avg. LT","Model-avg. UT","Model-avg. LT minus UT","cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$")
res<-res[,-1]
knitr::kable(res,"latex", booktabs = T, linesep = "",digits=3, escape=F,longtable=T,
             caption.short="Summary results for copulas for winter temperature versus green spruce aphid abundance",
      caption = "Summary results for analysis of copulas for winter temperature versus green spruce 
aphid abundance for each of the 10 sampling locations. The column labeled p, CvM is the p-value 
result for the goodness of fit test using the Cramer-von Mises statistic (see section on 
methods for the observational hypothesis). The 
column labeled p, KS is the p-value 
result for the goodness of fit test using the Kolmogorov-Smirnov statistic. Entries with an NA are because model selection and other statistics
were only employed for copulas for which independence was rejected (5 percent significance level).\\label{tab:greenspruce}") %>% column_spec(1:12,width="1.2 cm")#%>%
#kable_styling(latex_options = c("scale_down"))
```
